### 2012-11-19 doc/levelset/level_set_method_on_quadtree_JCP_rebuttal.tex

Table \ref{tab_circle_motion_by_curvature} demonstrates that the method is
first order accurate in the average norm for smooth a interface. The
deterioration in the maximum norm probably comes from the Elliptic part of
the solver, which propagates the errors from the regions where the grid
cells are coarse and unrefined to the regions where the grid cells are
refined. Figure \ref{fig_curvature_flow_in_octree} illustrates the motion
of an interface under mean curvature for the example presented in
\cite{Smereka:2003:Curvature}.

%So it is a first order discretization. One can go further for a better accuracy. However there is
%the following issue in this discretization. $\phi^{n+1}$ is calculated by solving a Poisson-type
%equation. As discussed in \cite{Min::Poisson_AMR_Second_Order}, the Poisson solver is second order
%accurate with the uniform refinement, i.e. splitting every cell. In level set method, it is natural
%to put the most resource near interface. So that in solving the Poisson-type equation, the errors
%in the coarse grids, i.e. the region far away from interface contaminates the region near
%interface, since the solution of one point affects the whole domain in Poisson equation. The
%accuracy of the Poisson solver in \cite{Min::Poisson_AMR_Second_Order} has not been analyzed in the
%adaptive refinement.

%\begin{figure}
%\begin{center}
%\includegraphics[scale=0.25]{curvature_flow_quadtree.eps}
%\end{center}
%\caption{$\alpha=1,\beta=1$ in $256^3$ resolution. From top-left to bottom-right, the times are $0$, $0.039$, %$0.107$, $0.507$, $0.800$ and $1.03$. The CFL condition is $\Delta t=\frac{\Delta x_{\text{smallest}}}2$. %Note that the velocities of convection and curvature cancel out not to move the side parts of the two circles %for some small time. }
%\label{fig_curvature_flow_in_quadtree}
%\end{figure}

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|}\hline
\multirow{2}{*}{} & \multicolumn{7}{|c|}{Finest Resolution}\\\cline{2-8}
                  & $128^2$ & rate & $256^2$ &rate & $512^2$ & rate &$1024^2$ \\\hline
$L_1$      error of $\phi$&$5.22\TT{-3}$&1.00&$2.60\TT{-3}$&0.96&$1.33\TT{-3}$&0.95&$6.95\TT{-4}$\\\hline
$L_\infty$ error of $\phi$&$5.47\TT{-3}$&0.93&$2.86\TT{-3}$&0.86&$1.56\TT{-3}$&0.81&$8.91\TT{-4}$\\\hline
\end{tabular}
\end{center}
\caption{Convergence rate for a circle with curvature dependent speed of
$\alpha=1.5$ and $\beta=1$. Initially circle is centered at $(0,0)$ with
radius one in a domain of $[-2,2]^2$. Test was run until $0.5$.The radius
$r(t)$ of the circle satisfies $r'=\alpha-\frac\beta{r}$ with $r(0)=1$.
$r(0.5)$ is approximated as $1.3108122$ from the ordinary differential
equation within error bound of $10^{-7}$.}
\label{tab_circle_motion_by_curvature}
\end{table}

%\begin{figure}
%\begin{center}
%\includegraphics[scale=0.25]{curvature_flow_quadtree.eps}
%\end{center}
%\caption{$\alpha=1,\beta=1$ in $256^3$ resolution. From top-left to
%bottom-right, the times are $0$, $0.039$, $0.107$, $0.507$, $0.800$ and
%$1.03$. The time step restriction is $\Delta t=\Delta
%x_{\text{smallest}}/2$. Note that the velocities of convection and
%curvature cancel out not to move the side parts of the two circles for some
%small time. } \label{fig_curvature_flow_in_quadtree}
%\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.25]{curvature_flow_in_octree.eps}
\end{center}
\caption{Motion with curvature flow for a barbel shape. $\alpha=0,\beta=1$ in $256^3$ resolution. From top-left to bottom-right, the times are $0$, $0.023$, $0.093$, $0.140$, $0.304$ and $0.323$. The CFL condition is $\Delta t=\Delta x_{\text{smallest}}$.}
\label{fig_curvature_flow_in_octree}
\end{figure}

%--------------------------------------------------------------------------
%
%
%
%--------------------------------------------------------------------------
\section{Adaptive Grid Generation}
As the interface deforms some provisions must be given to refine the grid
near the interface while coarsening in regions farther away. The grid is
constructed in such a way that the smallest grid cells lie on the interface
as described in section \ref{sec_spatial_discretization}. This construction
depends on an input function,
$\tilde{\phi}^{n+1}:\mathbb{R}^n\to\mathbb{R}$ that is close to the signed
distance function at any point in space. This function can be constructed
in two different ways: (1) In the case where semi-Lagrangian methods are
used, a function $\tilde{\phi}^{n+1}$ can be defined as
$\tilde{\phi}^{n+1}=\phi^n(x_d)$, where $x_d$ is found by tracing back the
characteristic curves and where $\phi^n(x_d)$ is interpolated from the node
values of $\phi^n$ as described in section \ref{sec_semi_lagrangian}. (2)
In the case where the velocity field is nonlinear, semi-Lagrangian methods
cannot be used. In this case the level set function is first evolved from
$\phi^{n}$ to $\phi^{n+1}$ on the same grid $G^n$. Then $\phi^{n+1}$ is
reinitialized into a signed distance function using the algorithm described
in section \ref{sec_reinitialization}. Now at every point in space, we can
define $\tilde{\phi}^{n+1}:\mathbb{R}^n\to\mathbb{R}$ by interpolation of
$\phi^{n+1}$. Once the function
$\tilde{\phi}^{n+1}:\mathbb{R}^n\to\mathbb{R}$ can be define anywhere in
space, the new grid $G^{n+1}$ is generated by simply splitting a cell if
the Lipschitz condition: $$\min_{v\in \textrm{vertices}(C)}|\phi(v)| \le
\textrm{Lip}(\phi)\cdot \textrm{diag-size}(C)$$ is satisfied. In practice,
instead of generating $G^{n+1}$ from the root cell, we start from $G^n$ and
apply the procedure detailed in Algorithm 1, i.e. starting the recursion
from the root cell of $G^{n+1}$, the cell is recursively split if the
refinement criteria is satisfied, otherwise all of its children are merged.

\begin{center}
\begin{tabular}{ll}\hline
\multicolumn{2}{c}{\textbf{Algorithm 1} : Grid Generation} \bigstrut\\\hline
\multicolumn{2}{l}{\textbf{Input} : $G^n$ and ${\tilde{\phi}}^{n+1}:\mathbb{R}^d \to \mathbb{R}$
}\bigstrut\\\hline
1. &$G^{n+1}=G^{n}$ \\
2. &$C= \text{the root cell of } G^{n+1}$ \\
3. &\textbf{if} the Lipschitz condition for  ${\tilde{\phi}}^{n+1}$ is satisfied at $C$ \\
4. &\qquad \textbf{if} $C$ is a leaf cell \\
5. &\qquad\qquad split $C$ \\
6. &\qquad \textbf{end if} \\
7. &\qquad \textbf{for} each child cell $C'$ of $C$ \\
8. &\qquad \qquad \textbf{go to} 3 with $C=C'$ \\
9. &\qquad \textbf{end for} \\
10.&\textbf{else} \\
11.&\qquad merge $C$ \\
12.& \textbf{end if} \\\hline
\multicolumn{2}{l}{\textbf{Output} : $G^{n+1}$}\bigstrut\\\hline
\end{tabular}
\end{center}

%-------------------------------------------------------------------------------
%
% Extrapolation
%
%-------------------------------------------------------------------------------
\section{High Order Extrapolation in the Normal Direction}
The ghost fluid method, introduced by Fedkiw \etal \cite{Fedkiw:1999:ANE}, is a technique for
imposing boundary conditions at the interface in a level set framework and has been successfully
applied to a wide range of applications (see e.g.
\cite{fedkiw:2001:VSO,fedkiw:2002:CEF,Fedkiw:Ghost_Fluid:Godunov_Methods,nguyen:2001:jcpflame,
Nguyen:2002:Conservative_Ghost_Fluid,Caiden:Incompressible_Compressible,Liu:2000:ABC,Gibou:2002:Second_Order_Accurate,Gibou:2003:Crystal_Growth_JSC,
Gibou:2003:Fourth_Order_Accurate,Gibou::Vaporization} and the references therein). One basic
component of this method is the extrapolation of some scalar quantities in the normal direction. In
some cases (see e.g. \cite{Gibou:2003:Fourth_Order_Accurate}), high order extrapolations in the
normal direction are needed. This can be performed in a series of steps, as proposed in Aslam
\cite{Aslam:Extrapolation}. For example, suppose that we seek to extrapolate $u$ quadratically from
the region where $\phi\leq 0$ to the region where $\phi>0$. We first compute
$u_{nn}=\vec{n}\cdot\nabla\left( \vec{n}\cdot\nabla u \right)$ in the region where $\phi \leq 0$
and extrapolate (constant extrapolation) this quantity across the interface by solving the
following partial differential equation:
\begin{equation*}
\frac{\partial u_{nn}}{\partial \tau} + H(\phi,u_{nn})(\vec{n}\cdot\nabla u_{nn}       )=0,
\end{equation*}
where $H(\phi,u_{nn})$ is the Heaviside function defined below and $\tau$ is the fictitious time
step. Then we define $u_n$ in the region where $\phi>0$ in such a way its normal derivative is
$u_{nn}$. This can be accomplished by solving the following PDE:
\begin{equation*}
\frac{\partial u_{n }}{\partial \tau} + H(\phi,u_n)(\vec{n}\cdot\nabla u_{n }-u_{nn})=0.
\end{equation*}
Finally we can define $u$ in such a way its normal derivative is $u_n$ by solving:
\begin{equation*}
\frac{\partial u_{  }}{\partial \tau} + H(\phi,u)(\vec{n}\cdot\nabla u     -u_{n })=0.
\end{equation*}
Numerically the Heaviside function $H(\phi,S)(v_i)$ associated with a quantity $S$ at the node
$v_i$ is set to zero if the nodes involved in the computation of $S$ are all in the region where
$\phi<0$. Otherwise, it is set to 1. Therefore we define the Heaviside functions $H(\phi,u)$,
$H(\phi,u_{n})$ and $H(\phi,u_{nn})$ as follows:
\begin{align*}
H(\phi,u)(v_i) &= \begin{cases}
                 0 & \text{ if $\phi(v_i)<0$ } \\
                 1 & \text{ otherwise}
                 \end{cases} ,\\
H(\phi,u_n)(v_i) &= \begin{cases}
                   0 & \text{ if $H(\phi,u)(v_j)=0$ for all $v_j\in$ ngbd$(v_i)$} \\
                   1 & \text{ otherwise}
                   \end{cases} ,\\
H(\phi,u_{nn})(v_i) &= \begin{cases}
                   0 & \text{ if $H(\phi,u_n)(v_j)=0$ for all $v_j\in$ ngbd$(v_i)$} \\
                   1 & \text{ otherwise}
                   \end{cases},
\end{align*}
where $\text{ngbd}(v_i)$ denotes the set of direct neighboring nodes of $v_i$. The quantity
$u_n=\vec{n}\cdot\nabla u$ is computed by the central finite differences described in section
\ref{sec_finite_differences} for all the nodes where $H(\phi,u_n)=0$. Likewise, using the values of
$u_{nn}$ is then computed by central differencing for all the nodes where $H(\phi,u_{nn})=0$. The
three partial differential equations above are discretized in a dimension by dimension framework
using the upwind schemes and the one-sided finite differences of section
\ref{sec_finite_differences}, i.e. the discretizations in a semi-discrete form read:
\begin{align*}
\frac{d}{d\tau}u_{nn}+H(\phi,u_{nn})\left(n_x^+D_x^-u_{nn}+n_x^-D_x^+u_{nn}\right)&=0,
\end{align*}
\begin{align*}
\frac{d}{d\tau}u_{n }+H(\phi,u_{n })\left(n_x^+D_x^-u_{n }+n_x^-D_x^+u_{n }\right)&=H(\phi,u_{n})
u_{nn},
\end{align*}
and
\begin{align*}
\frac{d}{d\tau}u     +H(\phi,u     )\left(n_x^+D_x^-u     + n_x^-D_x^+u    \right)&=H(\phi,u ) u_{n
}.
\end{align*}
These semi-discrete equations are then evolved in time using the same TVD RK-2 method of section
\ref{sec_reinitialization}. Since the equations are evolved in fictitious time, we can take the
same time step restriction as in the reinitialization procedure of section
\ref{sec_reinitialization}.

Figure \ref{extrapolation_aslam} illustrates the constant, linear and quadratic extrapolation
obtained with the algorithm described above: Consider a computational domain $\Omega=(-\pi,
\pi)\times(-\pi, \pi)$ separated into two regions: $\Omega^-$ defined as the interior of a disk
with center at the origin and radius two, and $\Omega^+=\Omega \setminus \Omega^-$. The function
$u$ to be extrapolated from $\Omega^-$ to $\Omega^+$ is defined as $u=\cos(x)\sin(y)$ for $x \in
\Omega^-$. We have extrapolated $u$ in the entire region in this example for the sake of
presentation but we emphasize that in practice the extrapolation is performed only in a
neighborhood of the interface. Tables \ref{table_constant_extrapolation},
\ref{table_linear_extrapolation} and \ref{table_quadratic_extrapolation} demonstrate the first
order accuracy for the constant extrapolation, the second order accuracy for the linear
extrapolation and the third order accuracy for the quadratic extrapolation. We note that it is
enough to discretize $D^{\pm}_xu_{nn}$ and $D^{\pm}_xu_{n}$ with the first order accurate finite
difference of section \ref{sec_finite_differences}, and $D^{\pm}_xu$ with the second order accurate
finite difference in section \ref{sec_finite_differences} to achieve third order accuracy in $u$ in
the case of a quadratic extrapolation. The same accuracy would be achieved in the case where the
second order accurate finite differences were used for $D^{\pm}_xu_{nn}$, $D^{\pm}_xu_{n}$, and
$D^{\pm}_xu_{n}$. However, using the first order accurate finite difference schemes for
$D^{\pm}_xu_{nn}$ and $D^{\pm}_xu_{n}$ yields more robust results since $u_{nn}$ and $u_{n}$ may be
noisy unless $u$ is a very smooth function.

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline Finest &$L^\infty$ error&rate&$L^1 error$&rate \\ Resolution& of $\phi$      &    & of
$\phi$ & \\\hline

$32^2$&$5.11\times10^{-1}$& &$1.11\times10^{-1}$&      \\\hline
$64^2$&$2.55\times10^{-1}$&1.01&$3.75\times10^{-2}$&1.56   \\\hline
$128^2$&$1.25\times10^{-1}$&1.02&$1.06\times10^{-2}$&1.82   \\\hline
$256^2$&$6.20\times10^{-2}$&1.01&$2.89\times10^{-3}$&1.87   \\\hline
$512^2$&$3.14\times10^{-2}$&.97&$7.59\times10^{-4}$&1.92   \\\hline
$1024^2$&$1.59\times10^{-2}$&.98&$2.01\times10^{-4}$&1.91   \\\hline
\end{tabular}
\end{center}
\caption{Convergence rate for the constant
extrapolation}\label{table_constant_extrapolation}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline Effective &$L^\infty$ error&rate&$L^1 error$&rate \\ Resolution& of $\phi$      &    & of
$\phi$ & \\\hline

$32^2$&$2.02\times10^{-1}$& &$3.52\times10^{-2}$&      \\\hline
$64^2$&$7.21\times10^{-2}$&1.48&$6.09\times10^{-3}$&2.53   \\\hline
$128^2$&$1.78\times10^{-2}$&2.00&$8.79\times10^{-4}$&2.79   \\\hline
$256^2$&$5.27\times10^{-3}$&1.76&$1.19\times10^{-4}$&2.88   \\\hline
$512^2$&$1.12\times10^{-3}$&2.22&$1.54\times10^{-5}$&2.94   \\\hline
$1024^2$&$2.83\times10^{-4}$&1.99&$2.04\times10^{-6}$&2.91   \\\hline
\end{tabular}
\end{center}
\caption{Convergence rate for the linear
extrapolation}\label{table_linear_extrapolation}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline Effective &$L^\infty$ error&rate&$L^1 error$&rate \\ Resolution& of $\phi$      &    & of
$\phi$ & \\\hline

$32^2$&$1.62\times10^{-1}$& &$2.26\times10^{-2}$&      \\\hline
$64^2$&$2.31\times10^{-2}$&2.82&$2.21\times10^{-3}$&3.36   \\\hline
$128^2$&$2.95\times10^{-3}$&2.96&$1.65\times10^{-4}$&3.73   \\\hline
$256^2$&$3.81\times10^{-4}$&2.95&$1.17\times10^{-5}$&3.81   \\\hline
$512^2$&$4.89\times10^{-5}$&2.96&$7.82\times10^{-7}$&3.91   \\\hline
$1024^2$&$6.19\times10^{-6}$&2.98&$5.31\times10^{-8}$&3.88   \\\hline
\end{tabular}
\end{center}
\caption{Convergence rate for the quadratic
extrapolation}\label{table_quadratic_extrapolation}
\end{table}

### 2012-11-19 doc/levelset/level_set_method_on_quadtree_JCP_rebuttal.tex

\begin{figure}
\begin{center}
\includegraphics[scale=0.25]{constant_extrapolation.eps}\includegraphics[scale=0.25]{linear_extrapolation.eps}\includegraphics[scale=0.25]{quadratic_extrapolation.eps}
\includegraphics[scale=0.2]{constant_extrapolation_zoom.eps}\hspace{1cm}\includegraphics[scale=0.2]{linear_extrapolation_zoom.eps}\hspace{1cm}\includegraphics[scale=0.2]{quadratic_extrapolation_zoom.eps}
\end{center}
\caption{Contours of the solution after it has been extrapolated across the
interface with a constant (left), linear (center) and quadratic (right)
extrapolations across an interface. The top row illustrates the
extrapolation on the entire domain and the bottom row is a zoom near the
interface. The exact solution is given inside the circle centered at the
origin and with radius 2 and is extrapolated outside in the normal
direction. We then plot the level curves of the solution.}
\label{extrapolation_aslam}
\end{figure}

\section{Conclusion}
We have presented a level set method on non-graded adaptive Cartesian
grids, i.e. grids for which the ratio between adjacent cells is not
constrained. We use quadtree and octree data structures to represent the
grid and a simple algorithm to generate a mesh with the finest resolution
at the interface. We have presented (1) a locally third order accurate
reinitialization scheme that transforms an arbitrary level set function
into a signed distance function, (2) a second order accurate
semi-Lagrangian methods to evolve the linear level set advection equation
under an externally generated velocity field, (3) a second order accurate
upwind method to evolve the nonlinear level set equation under a normal
velocity as well as to extrapolate scalar quantities across an interface in
the normal direction, and (4) a semi-implicit scheme to evolve the
interface under mean curvature. This method produces results with a
negligible amount of mass loss. We have proposed numerical examples in two
and three spatial dimensions to demonstrate the accuracy of the method.

\section{Acknowledgment}
The research of F. Gibou was supported in part by the Alfred P. Sloan Foundation through a research
fellowship in Mathematics.
%-------------------------------------------------------------------------------
%
% References : "reference.bib"
%
%-------------------------------------------------------------------------------
\bibliographystyle{plain}
\addcontentsline{toc}{section}{\refname}\bibliography{../references}

\end{document}
